--- 
title: "EECS E6690 hw3"
author: "Chong Hu ch3467"
date: 'Sep 22, 2019'
output: 
  pdf_document:
    latex_engine: pdflatex
header-includes: 
- \usepackage{graphicx}
- \usepackage{float}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.pos = 'H')
```

\section*{P1}
\begin{enumerate}
    \item[(a)] \begin{align*}
        \mathbb{P}[Y= \text{true}] &= \frac{\exp{(\hat{\beta}_0 + \hat{\beta}_1 X_1 + \hat{\beta}_2 X_2 )}}{1+ \exp{(\hat{\beta}_0 + \hat{\beta}_1 X_1 + \hat{\beta}_2 X_2 )}} \\
        &= \frac{\exp{(-6 + 0.05 * 40 + 1 * 3.5)}}{1 + \exp{(-6 + 0.05 * 40 + 1 * 3.5)}} \\
        & = 0.3775407
    \end{align*}
    Hence, $\mathbb{P}[Y= \text{true}] = 0.3775407$.
    \item[(b)]
    \begin{align*}
        \mathbb{P}[Y= \text{true}] = \frac{\exp{(\hat{\beta}_0 + \hat{\beta}_1 X_1 + \hat{\beta}_2 X_2 )}}{1+ \exp{(\hat{\beta}_0 + \hat{\beta}_1 X_1 + \hat{\beta}_2 X_2 )}}  &= 0.5\\
        \exp{(\hat{\beta}_0 + \hat{\beta}_1 X_1 + \hat{\beta}_2 X_2 )} &= 1\\
        -6 + 0.05 * x_1 + 1* 3.5 &= 0\\
        x_1 &= 50
    \end{align*}
     The student in part (a) need to study 50 hours to have a 50\% chance of getting an A in the class.
\end{enumerate}

\section*{P2}
\begin{align*}
    f_{\text{Yes}}(x) &= \frac{1}{\sqrt{2\pi \hat{\sigma} ^2}} \exp{- \frac{ (x - \mu_{\text{Yes}})^2}{2\hat{\sigma} ^2 }}\\
    f_{\text{No}}(x) &= \frac{1}{\sqrt{2\pi \hat{\sigma} ^2}} \exp{- \frac{ (x - \mu_{\text{No}})^2}{2\hat{\sigma} ^2 }}\\
    \mathbb{P}[Y = \text{Yes} \mid X = 4 ] &= \frac{\pi_{\text{Yes}} f_{\text{Yes}}(x)}{\pi_{\text{Yes}} f_{\text{Yes}}(x) + \pi_{\text{No}} f_{\text{No}}(x)}\\
    &= \frac{0.8 * 0.04032845}{0.8* 0.04032845 + 0.2 * 0.05324133 }\\
    &= 0.7518524
\end{align*}

\section*{P3}
\begin{align*}
    \text{likelihood} &= \prod_{i = 1} ^ n \bigg( \frac{\exp{(\beta_0 + \beta_1 x_i)}}{1+ \exp{(\beta_0 + \beta_1 x_i)}} \bigg)^{y_i} \bigg( \frac{1}{1+ \exp{(\beta_0 + \beta_1 x_i)}} \bigg)^{1- y_i}  \\
    l(\beta_0, \beta_1) &= \sum_{i = 1}^{n} y_i (\beta_0 + \beta_1 x_i) - \ln{\big( 1+ \exp{(\beta_0 + \beta_1 x_i)} \big)}
\end{align*}
To maximize the likelihood, we need to set
\begin{align*}
    \frac{\partial  l(\beta_0, \beta_1)}{\partial \beta_0 } &= \sum_{i=1}^{n} \bigg( y_i - \frac{\exp{(\hat{ \beta}_0 + \hat{\beta}_1 x_i)}}{1+ \exp{( \hat{\beta}_0 + \hat{\beta}_1 x_i)}}\bigg) = 0 \\
    \frac{\partial  l(\beta_0, \beta_1)}{\partial \beta_1 } &= \sum_{i=1}^{n} x_i \bigg( y_i - \frac{\exp{(\hat{ \beta}_0 + \hat{\beta}_1 x_i)}}{1+ \exp{( \hat{\beta}_0 + \hat{\beta}_1 x_i)}}\bigg) = 0
\end{align*}

Using the Newton–Raphson algorithm, I write all variables in matrix form. $\mathbf{W}$ is $n \times n$ diagonal matrix of weights with $i$th diagonal element $p(x;\beta^{\text{old}}) (1 - p(x; \beta^{\text{old}}))$.
\begin{align*}
    \mathbf{z} & = \mathbf{X} \beta^{\text{old}} + \mathbf{W} ^{-1} (\mathbf{y} - \mathbf{p})\\
    \beta^{\text{new}} &= (\mathbf{X}^T \mathbf{WX})^ {-1} \mathbf{X}^T \mathbf{W} \mathbf{z} 
\end{align*}
Then we start with $\beta_0 = 0$ and $\beta_1 = 0$ and perform 10 iterations.
```{r}
library(matlib)
```
```{r}
x = c(0.0, 0.2, 0.4, 0.6, 0.8 , 1.0)
X = matrix(cbind(1, x), ncol = 2) 
Y = matrix(c(0, 0, 0, 1, 0, 1), ncol = 1)
beta_old = matrix(c(0, 0), ncol = 1)
beta_new = beta_old
for(i in 1:10){
  p = exp(X %*% beta_old) / (1 + exp(X %*% beta_old))
  w = as.vector(p * (1 - p))
  W = diag(w)
  z = X %*% beta_old + inv(W) %*% (Y - p)
  beta_new = inv(t(X)  %*% W %*% X) %*% t(X) %*% W %*% z
  beta_old = beta_new
}
beta_new
```
After 10 iterations, we can get $\hat{\beta}_0 =$ `r beta_new[1]` and $\hat{\beta}_1 =$ `r beta_new[2]`.

\section*{P4}
\begin{align*}
    \mathrm{Cov}[\mathbf{Y}] &= \mathrm{Cov}[\mathbf{AX}]\\
    &= \mathbf{A} \text{Cov}[\mathbf{X}] \mathbf{A}^T\\
    &= \mathbf{A} \mathbf{\Sigma} \mathbf{A}^T
\end{align*}
Then, since $\text{cov}[\mathbf{Y}]$ is an identity matrix, we have 
\begin{align*}
    \mathbf{A} \mathbf{\Sigma} \mathbf{A}^T &= \mathbf{I}
\end{align*}
Since $\mathbf{\Sigma}$ is symmetric, we can get
\begin{equation*}
    \mathbf{\Sigma} = \mathbf{Q} \mathbf{\Lambda} \mathbf{Q}^T
\end{equation*}
where $\mathbf{Q}$ is an orthogonal matrix whose columns are the eigenvectors of $\mathbf{\Sigma}$, and $\Lambda$ is a diagonal matrix whose entries are the eigenvalues of $\mathbf{\Sigma}$. 
in
And,
\begin{equation*}
    \mathbf{Q} \mathbf{Q}^T = \mathbf{I}
\end{equation*}
Then,
\begin{align*}
    \mathbf{A} \mathbf{Q} \mathbf{\Lambda} \mathbf{Q}^T \mathbf{A}^T &= \mathbf{I} \\
    (\mathbf{A} \mathbf{Q}) \mathbf{\Lambda} (\mathbf{A} \mathbf{Q} )^T &= \mathbf{I}\\
    \mathbf{A} \mathbf{Q} &= (\mathbf{\Lambda}^{-1})^{\frac{1}{2}} \\
    \mathbf{A} &=  (\mathbf{\Lambda}^{-1})^{\frac{1}{2}} \mathbf{Q}^{-1} \\
\end{align*}
\begin{align*}
        \mathbf{\Lambda} &=  
  \left[ {\begin{array}{cc}
  \sigma_1^2/2 - (4*\rho^2\sigma_1^2\sigma_2^2 + \sigma_1^4 - 2\sigma_1^2\sigma_2^2 + \sigma_2^4)^{1/2}/2 + \sigma_2^2/2 &  \\
  0 & \\
  \end{array} }\right. \\
  &
  \left. {\begin{array}{cc}
  & 0
     \\
      & (4\rho^2\sigma_1^2\sigma_2^2 + \sigma_1^4 - 2\sigma_1^2\sigma_2^2 + \sigma_2^4)^{1/2}/2 + \sigma_1^2/2 + \sigma_2^2/2 \\
  \end{array}} \right] \\
  \mathbf{Q} &=  
  \left[ {\begin{array}{cc}
  (\sigma_1^2/2 - (4\rho^2\sigma_1^2\sigma_2^2 + \sigma_1^4 - 2\sigma_1^2\sigma_2^2 + \sigma_2^4)^{1/2}/2 + \sigma_2^2/2)/(\rho\sigma_1\sigma_2) - \sigma_2/(\rho\sigma_1) &  \\
  1 & \\
  \end{array} }\right. \\
  &
  \left. {\begin{array}{cc}
  & ((4\rho^2\sigma_1^2\sigma_2^2 + \sigma_1^4 - 2\sigma_1^2\sigma_2^2 + \sigma_2^4)^{1/2}/2 + \sigma_1^2/2 + \sigma_2^2/2)/(\rho\sigma_1\sigma_2) - \sigma_2/(\rho\sigma_1) 
     \\
      & 1 \\
  \end{array}} \right]
\end{align*}
Therefore,
\begin{align*}
  \mathbf{A} &=  
  \left[ {\begin{array}{cc}
- \frac{ (2^{1/2}\rho\sigma_1\sigma_2(1/(\sigma_1^2 - (4*\rho^2*\sigma_1^2\sigma_2^2 + \sigma_1^4 - 2\sigma_1^2\sigma_2^2 + \sigma_2^4)^{1/2} + \sigma_2^2))^{1/2})}{(4*\rho^2\sigma_1^2\sigma_2^2 + \sigma_1^4 - 2\sigma_1^2\sigma_2^2 + \sigma_2^4)^{1/2}} &  \\
  \frac{2^{1/2}\rho\sigma_1\sigma_2(1/((4\rho^2\sigma_1^2\sigma_2^2 + \sigma_1^4 - 2\sigma_1^2\sigma_2^2 + \sigma_2^4)^{1/2} + \sigma_1^2 + \sigma_2^2))^{1/2})}{(4\rho^2\sigma_1^2\sigma_2^2 + \sigma_1^4 - 2\sigma_1^2\sigma_2^2 + \sigma_2^4)^{1/2}} & \\
  \end{array} }\right. \\
  &
  \left. {\begin{array}{cc}
  & \frac{ (2^{1/2}*(1/(\sigma_1^2 - (4\rho^2\sigma_1^2\sigma_2^2 + \sigma_1^4 - 2\sigma_1^2\sigma_2^2 + \sigma_2^4)^(1/2) + \sigma_2^2))^{1/2}((4\rho^2\sigma_1^2\sigma_2^2 + \sigma_1^4 - 2\sigma_1^2\sigma_2^2 + \sigma_2^4)^{1/2} + \sigma_1^2 - \sigma_2^2))}{(2(4\rho^2\sigma_1^2\sigma_2^2 + \sigma_1^4 - 2\sigma_1^2\sigma_2^2 + \sigma_2^4)^{1/2})}
     \\
      & \frac{(2^{1/2}*(1/((4\rho^2\sigma_1^2\sigma_2^2 + \sigma_1^4 - 2\sigma_1^2\sigma_2^2 + \sigma_2^4)^{1/2} + \sigma_1^2 + \sigma_2^2))^{1/2}((4\rho^2\sigma_1^2\sigma_2^2 + \sigma_1^4 - 2\sigma_1^2\sigma_2^2 + \sigma_2^4)^{1/2} - \sigma_1^2 + \sigma_2^2))}{(2(4\rho^2\sigma_1^2\sigma_2^2 + \sigma_1^4 - 2\sigma_1^2\sigma_2^2 + \sigma_2^4)^{1/2})} \\
  \end{array}} \right]
\end{align*}

\section*{P5}
Since
\begin{align*}
    \hat{\mu}_k = \frac{1}{n_k} \sum_{i:\ y_i = k} x_i \quad\text{and}\quad \hat{\sigma}_k^2 = \frac{1}{n_k-1} \sum_{i:\ y_i = k} (x_i -\hat{\mu}_k)^2
\end{align*}
and under the Gaussian assumption, the variance of each population is $\sigma^2$, we can easily get
 \begin{align*}
     \mathbb{E}[\hat{\sigma}_k^2] = \sigma^2
 \end{align*}
Hence, 
\begin{align*}
     \mathbb{E}[\hat{\sigma}^2] &= \mathbb{E}\bigg[\sum_{k=1}^K \alpha_k \hat{\sigma}_k^2 \bigg] = \sum_{k=1}^K \alpha_k\mathbb{E}\big[ \hat{\sigma}_k^2\big] \\
     &= \sigma^2 \sum_{k=1}^K \alpha_k = \sigma^2
\end{align*}
Also, under the Gaussian assumptions, we have
\begin{align*}
    \frac{(n_k-1) \hat{\sigma}_k^2}{\sigma^2} & \sim  \mathcal{X}_{n_k-1}^2 \\
    \mathrm{Var}\bigg[ \frac{(n_k-1) \hat{\sigma}_k^2}{\sigma^2} \bigg] &= \mathrm{Var} \big[\mathcal{X}_{n_k-1}^2\big] \\
    \frac{(n_k-1)^2 }{\sigma^4} \mathrm{Var} [\hat{\sigma}_k^2] &= 2(n_k-1)\\
    \mathrm{Var} [\hat{\sigma}_k^2] &= \frac{2 \sigma^4}{n_k-1}
\end{align*}


Then, 
\begin{align*}
    \mathrm{Var}\big[\hat{\sigma}^2 \big] &= \mathrm{Var}\bigg[\sum_{k=1}^K \alpha_k \hat{\sigma}_k^2 \bigg]\\
    &=\sum_{k=1}^K  \alpha_k^2 \mathrm{Var} \big[ \hat{\sigma}_k^2 \big]\\
    &= \sum_{k=1}^K  \alpha_k^2 \frac{2 \sigma^4}{n_k-1} \\
    &= 2 \sigma^4 \sum_{k=1}^K \frac{ \alpha_k^2}{n_k-1} \\
    &= 2 \sigma^4 \sum_{k=1}^K \frac{ \alpha_k^2}{n_k-1} * 1 \\
    &=  2 \sigma^4 \sum_{k=1}^K \frac{ \alpha_k^2}{n_k-1} * \sum_{k=1}^{K} \frac{n_k-1}{K-1} \\
    \text{where\ } &  \sum_{k=1}^K \alpha_k = 1 \\
    \text{and\ } &  \sum_{k=1}^K \frac{n_k-1}{n-K} = 1 \\
    &= \frac{2 \sigma^4}{n-K}  \sum_{k=1}^K \frac{ \alpha_k^2}{n_k-1} * \sum_{k=1}^{K}n_k-1
\end{align*} 
Based on \texttt{Cauchy-Schwarz\ inequality},
\begin{align*}
    \mathrm{Var}\big[\hat{\sigma}^4 \big] &=   \frac{2 \sigma^2}{n-K}  \sum_{k=1}^K \frac{ \alpha_k^2}{n_k-1} * \sum_{k=1}^{K}n_k-1\\
    & \geq \frac{2 \sigma^4}{n-K} \Bigg( \sum_{k=1}^K \sqrt{\frac{\alpha_k^2}{\sqrt{n_k-1}} * \sqrt{n_k-1}} \Bigg)^2\\
    &=  \frac{2 \sigma^4}{n-K} \bigg( \sum_{k=1}^K \alpha_k \bigg)\\
    &=  \frac{2 \sigma^4}{n-K} 
\end{align*}
where equality happens if and only if $\alpha_k = (n_k-1)/(n-K)$.
Therefore, the variance of $\hat{\sigma}^2$ is minimized when $\alpha_k = (n_k-1)/(n-K)$.

\section*{P6}
Through majority vote approach, six out of ten estimates are this class is Red. Therefore, the final classification is Red.
Through average probability, the average estimate is 0.45, therefore the final classification is Green.


\section*{P7}
```{r}
library(ISLR)
library(tree)
attach(OJ)
set.seed(1000)
```

\subsection*{(a)}
```{r}
train = sample(dim(OJ)[1], 800)
OJ.train = OJ[train, ]
OJ.test = OJ[-train, ]
```

\subsection*{(b)}
```{r}
mod.tr <- tree(Purchase ~ ., data = OJ.train)
summary(mod.tr)
```
Training error rate for the tree is 0.16. The tree only use three variables `LoyalCH`, `PriceDiff` and `SalePriceMM`. There are 8 terminal nodes in the tree.

## (c)
```{r}
mod.tr
```

Let's interpret the terminal node with "10)". The splitting variable of this node is `PriceDiff`. `PriceDiff < 0.05` will be predicted as `MM` for the response `Purchase`. There are 78 points in total below this node. The deviation of those points is 79.16.  0.20513% points  have `CH` as value for `Purchase` and 0.79487% points have `MM` as value for `purchase`. The * in the end means this is a terminal node.

## (d)

```{r, fig.height = 4.5, fig.width = 7, fig.align = "center",  fig.cap="Plot of the tree"}
plot(mod.tr)
text(mod.tr, pretty = 5)
```
Three split data points into two sets based on `LoyalCH < 0.5036` at the root.Then, the subtrees split data points into different leaves based on different conditions. Among those conditions, we can clearly see that variable `LoyalCH` is used very often. In the end, we get eight terminal nodes to predict response `Purchase` as `CH` or `MM`.

## (e)

```{r}
mod.pred = predict(mod.tr, OJ.test, type = "class")
err_rate = sum(OJ.test$Purchase != mod.pred) / dim(OJ.test)[1]
print(sprintf("err rate: %.2f%% ",  err_rate*100))
table(OJ.test$Purchase, mod.pred)
```
Error rate on test set is `r sprintf("%.2f%%",  err_rate*100)`.

## (f)
```{r}
tree.cv = cv.tree(mod.tr, FUN = prune.tree)
tree.cv
```
The optimal size is 8 based on deviation.

## (g)
```{r, fig.height = 4.5, fig.width = 6, fig.align = "center",  fig.cap="cross-validated classification error vs. Tree size"}
plot(tree.cv$size, tree.cv$dev, col = "blue", pch = 3, 
     type = "b", xlab = "Tree Size", ylab = "cross-validated classification error")
```

## (h)
When the tree size is 8, cross-validation error is the lowest.

## (i)
```{r}
tree.pruned = prune.tree(mod.tr, best = 8)
tree.pruned
```
```{r}
#  create a pruned tree with five terminal nodes.
tree.pruned = prune.tree(mod.tr, best = 5)
tree.pruned
```

## (j)
```{r}
summary(tree.pruned)
```
The error rate of pruned tree on training set is 19.62% while the error rate of unpruned tree is 16%. After pruning, the error rate on training set is becoming higher. This is because we are removing some nodes in the tree and  reducing the overfitting on trainning set.

## (k)

```{r}
pruned.pred = predict(tree.pruned, OJ.test, type = "class")
err_rate = sum(OJ.test$Purchase != pruned.pred) / dim(OJ.test)[1]
print(sprintf("err rate: %.2f%% ",  err_rate*100))
```
The error rate of pruned tree on training set is 21.11% while the error rate of unpruned tree is 18.15%. In this scenario, the error rate of pruned tree is higher.

