--- 
title: "EECS E6690 hw3"
author: "Chong Hu ch3467"
date: 'Sep 22, 2019'
output: 
  pdf_document:
    latex_engine: pdflatex
header-includes: 
- \usepackage{graphicx}
- \usepackage{float}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.pos = 'H')
```

\section*{P1}
\begin{enumerate}
    \item[(a)] \begin{align*}
        \mathbb{P}[Y= \text{true}] &= \frac{\exp{(\hat{\beta}_0 + \hat{\beta}_1 X_1 + \hat{\beta}_2 X_2 )}}{1+ \exp{(\hat{\beta}_0 + \hat{\beta}_1 X_1 + \hat{\beta}_2 X_2 )}} \\
        &= \frac{\exp{(-6 + 0.05 * 40 + 1 * 3.5)}}{1 + \exp{(-6 + 0.05 * 40 + 1 * 3.5)}} \\
        & = 0.3775407
    \end{align*}
    Hence, $\mathbb{P}[Y= \text{true}] = 0.3775407$.
    \item[(b)]
    \begin{align*}
        \mathbb{P}[Y= \text{true}] = \frac{\exp{(\hat{\beta}_0 + \hat{\beta}_1 X_1 + \hat{\beta}_2 X_2 )}}{1+ \exp{(\hat{\beta}_0 + \hat{\beta}_1 X_1 + \hat{\beta}_2 X_2 )}}  &= 0.5\\
        \exp{(\hat{\beta}_0 + \hat{\beta}_1 X_1 + \hat{\beta}_2 X_2 )} &= 1\\
        -6 + 0.05 * x_1 + 1* 3.5 &= 0\\
        x_1 &= 50
    \end{align*}
     The student in part (a) need to study 50 hours to have a 50\% chance of getting an A in the class.
\end{enumerate}

\section*{P2}
\begin{align*}
    f_{\text{Yes}}(x) &= \frac{1}{\sqrt{2\pi \hat{\sigma} ^2}} \exp{- \frac{ (x - \mu_{\text{Yes}})^2}{2\hat{\sigma} ^2 }}\\
    f_{\text{No}}(x) &= \frac{1}{\sqrt{2\pi \hat{\sigma} ^2}} \exp{- \frac{ (x - \mu_{\text{No}})^2}{2\hat{\sigma} ^2 }}\\
    \mathbb{P}[Y = \text{Yes} \mid X = 4 ] &= \frac{\pi_{\text{Yes}} f_{\text{Yes}}(x)}{\pi_{\text{Yes}} f_{\text{Yes}}(x) + \pi_{\text{No}} f_{\text{No}}(x)}\\
    &= \frac{0.8 * 0.04032845}{0.8* 0.04032845 + 0.2 * 0.05324133 }\\
    &= 0.7518524
\end{align*}

\section*{P3}
\begin{align*}
    \text{likelihood} &= \prod_{i = 1} ^ n \bigg( \frac{\exp{(\beta_0 + \beta_1 x_i)}}{1+ \exp{(\beta_0 + \beta_1 x_i)}} \bigg)^{y_i} \bigg( \frac{1}{1+ \exp{(\beta_0 + \beta_1 x_i)}} \bigg)^{1- y_i}  \\
    l(\beta_0, \beta_1) &= \sum_{i = 1}^{n} y_i (\beta_0 + \beta_1 x_i) - \ln{\big( 1+ \exp{(\beta_0 + \beta_1 x_i)} \big)}
\end{align*}
To maximize the likelihood, we need to set
\begin{align*}
    \frac{\partial  l(\beta_0, \beta_1)}{\partial \beta_0 } &= \sum_{i=1}^{n} \bigg( y_i - \frac{\exp{(\hat{ \beta}_0 + \hat{\beta}_1 x_i)}}{1+ \exp{( \hat{\beta}_0 + \hat{\beta}_1 x_i)}}\bigg) = 0 \\
    \frac{\partial  l(\beta_0, \beta_1)}{\partial \beta_0 } &= \sum_{i=1}^{n} x_i \bigg( y_i - \frac{\exp{(\hat{ \beta}_0 + \hat{\beta}_1 x_i)}}{1+ \exp{( \hat{\beta}_0 + \hat{\beta}_1 x_i)}}\bigg) = 0
\end{align*}

Using the Newton–Raphson algorithm, I write all variables in matrix form. $\mathbf{W}$ is $n \times n$ diagonal matrix of weights with $i$th diagonal element $p(x;\beta^{\text{old}}) (1 - p(x; \beta^{\text{old}}))$.
\begin{align*}
    \mathbf{z} & = \mathbf{X} \beta^{\text{old}} + \mathbf{W} ^{-1} (\mathbf{y} - \mathbf{p})\\
    \beta^{\text{new}} &= (\mathbf{X}^T \mathbf{WX})^ {-1} \mathbf{X}^T \mathbf{W} \mathbf{z} 
\end{align*}
Then we start with $\beta_0 = 0$ and $\beta_1 = 0$ and perform 10 iterations.
```{r}
library(matlib)
```
```{r}
x = c(0.0, 0.2, 0.4, 0.6, 0.8 , 1.0)
X = matrix(cbind(1, x), ncol = 2) 
Y = matrix(c(0, 0, 0, 1, 0, 1), ncol = 1)
beta_old = matrix(c(0, 0), ncol = 1)
beta_new = beta_old
for(i in 1:10){
  p = exp(X %*% beta_old) / (1 + exp(X %*% beta_old))
  w = as.vector(p * (1 - p))
  W = diag(w)
  z = X %*% beta_old + inv(W) %*% (Y - p)
  beta_new = inv(t(X)  %*% W %*% X) %*% t(X) %*% W %*% z
  beta_old = beta_new
}
beta_new
```
After 10 iterations, we can get $\hat{\beta}_0 =$ `r beta_new[1]` and $\hat{\beta}_1 =$ `r beta_new[2]`.

\section*{P4}
\section*{P5}
\section*{P6}


\section*{P7}
```{r}
library(ISLR)
library(tree)
attach(OJ)
set.seed(1000)
```

\subsection*{(a)}
```{r}
train = sample(dim(OJ)[1], 800)
OJ.train = OJ[train, ]
OJ.test = OJ[-train, ]
```

\subsection*{(b)}
```{r}
mod.tr <- tree(Purchase ~ ., data = OJ.train)
summary(mod.tr)
```
Training error rate for the tree is 0.16. The tree only use three variables `LoyalCH`, `PriceDiff` and `SalePriceMM`. There are 8 terminal nodes in the tree.

## (c)
```{r}
mod.tr
```

Let's interpret the terminal node with "10)". The splitting variable of this node is `PriceDiff`. `PriceDiff < 0.05` will be predicted as `MM` for the response `Purchase`. There are 78 points in total below this node. The deviation of those points is 79.16.  0.20513% points  have `CH` as value for `Purchase` and 0.79487% points have `MM` as value for `purchase`. The * in the end means this is a terminal node.

## (d)

```{r, fig.height = 4.5, fig.width = 7, fig.align = "center",  fig.cap="Plot of the tree"}
plot(mod.tr)
text(mod.tr, pretty = 5)
```
Three split data points into two sets based on `LoyalCH < 0.5036` at the root.Then, the subtrees split data points into different leaves based on different conditions. Among those conditions, we can clearly see that variable `LoyalCH` is used very often. In the end, we get eight terminal nodes to predict response `Purchase` as `CH` or `MM`.

## (e)

```{r}
mod.pred = predict(mod.tr, OJ.test, type = "class")
err_rate = sum(OJ.test$Purchase != mod.pred) / dim(OJ.test)[1]
print(sprintf("err rate: %.2f%% ",  err_rate*100))
table(OJ.test$Purchase, mod.pred)
```
Error rate on test set is `r sprintf("%.2f%%",  err_rate*100)`.

## (f)
```{r}
tree.cv = cv.tree(mod.tr, FUN = prune.tree)
tree.cv
```
The optimal size is 8 based on deviation.

## (g)
```{r, fig.height = 4.5, fig.width = 6, fig.align = "center",  fig.cap="cross-validated classification error vs. Tree size"}
plot(tree.cv$size, tree.cv$dev, col = "blue", pch = 3, 
     type = "b", xlab = "Tree Size", ylab = "cross-validated classification error")
```

## (h)
When the tree size is 8, cross-validation error is the lowest.

## (i)
```{r}
tree.pruned = prune.tree(mod.tr, best = 8)
tree.pruned
```
```{r}
#  create a pruned tree with five terminal nodes.
tree.pruned = prune.tree(mod.tr, best = 5)
tree.pruned
```

## (j)
```{r}
summary(tree.pruned)
```
The error rate of pruned tree on training set is 19.62% while the error rate of unpruned tree is 16%. After pruning, the error rate on training set is becoming higher. This is because we are removing some nodes in the tree and  reducing the overfitting on trainning set.

## (k)

```{r}
pruned.pred = predict(tree.pruned, OJ.test, type = "class")
err_rate = sum(OJ.test$Purchase != pruned.pred) / dim(OJ.test)[1]
print(sprintf("err rate: %.2f%% ",  err_rate*100))
```
The error rate of pruned tree on training set is 21.11% while the error rate of unpruned tree is 18.15%. In this scenario, the error rate of pruned tree is higher.

